Histogram of Gradient Orientations of Signal Plots applied to P300 Detection

### 2017-12-21 article.tex
Casi listo

% Keywords
\keyword{EEG; BCI; P300; ALS; NBNN; HGOSP; SIFT}

@@

\begin{equation}
z(c) = \left \lfloor{ \frac{\max_{1 \leq n\leq n_{max}} \tilde{x}(n,c)  - \min_{1 \leq n\leq n_{max}} \tilde{x}(n,c) }{2} }\right \rfloor -   \left \lfloor{ \frac{\max_{1 \leq n\leq n_{max}} \tilde{x}(t,c)  + \min_{1 \leq n\leq n_{max}} \tilde{x}(n,c)}{ 2} }\right \rfloor
\label{eq:zerolevel}
\end{equation}

@@

On the generated image $I$, a keypoint $\mathbf{kp}$ is placed on a pixel $(x_{kp}, y_{kp})$ over the image plot and a window around the keypoint is considered. A local image patch of size $S_p \times S_p$ pixels is constructed by dividing the window in $16$ blocks of size $3s$ each one,  where $s$ is the scale of the local patch and it is an input parameter of the algorithm. It is arranged in a $4 \times 4$ grid and the pixel $ \mathbf{kp}$ is the patch center, thus $S_p = 4.3s $ pixels.

@@

\noindent  where $\mathbf{p}$ is a pixel from  the $i,j$-block $B_{i,j}$, $ i,j \in \{1,2,3,4\} $,  $\theta$ is the angle bin with $ \theta \in \{0, 45, 90, 135, 180, 225, 270, 315\} $,  $ |J(\mathbf{p})| $ is the norm of the gradient vector in the pixel $\mathbf{p}$ and it is computed using finite differences, $\angle J(\mathbf{p}) $ is the angle of the gradient vector, and $ w_\mathrm{ang}(\cdot) $  and $ w_{ij}(\cdot) $ are linear interpolation functions used by Lowe and Vedaldi et al. in~\citep{Lowe2004,Vedaldi2010}.  Lastly, the fixed value of $ 3 $ is a magnification factor which corresponds to the number of pixels per each block when $s = 1$.  As the patch has  $16$ blocks and  $8$ bin angles are considered, a descriptor of $128$ dimension is obtained. It can be observed that in each step, the histogram is computed by multiplying by $ |J(\mathbf{p})| $, so the method considers both, the magnitude and the orientation of the gradient vector.

@@

\begin{enumerate}
\item \label{paso1}Highlight randomly the rows and columns from the matrix.  There is one row and one column that should match the letter selected by the subject.
\item  \label{paso2} Repeat step~\ref{paso1} $k_a$ times, obtaining the single trial segments $S_1(n,c),\dots,S_{k_a}(n,c)$, where the variables $n \in \{1, \dots, n_{max}\}$ and $c \in \{1,2,\dots,Ch\}$ correspond to sample points and channel, respectively. The parameter $k_a$ is the number of repetitions and it is an input parameter of the algorithm.
\item \label{paso3} Compute the Ensemble Average by
\begin{equation}
x(n,c)= \frac{1}{k_a}\sum_{i=1}^{k_a}S_i(n,c),n \in \{1, \dots, n_{max}\}, c \in \{1,\dots,Ch\}
\label{averaging}
\end{equation}  
for each row and column.

\item \label{paso4}Plot the signal $x(n,c)$,  $n \in \{1, \dots, n_{max}\}$, $c \in \{1,\dots,Ch\}$,  according Section~\ref{Plot}.

@@

where $NN_T(d^l_u), \;l\in\{row,{col}\}$ is the set of the $k$ nearest neighbors to $d^l_u, \;l\in\{row,{col}\}$ and $q$ is a template descriptor that belongs to $NN_T(d^l_u), \;l\in\{row,{col}\}$.  This set is obtained according to the k-NNBN  algorithm~\citep{Boiman2008}.

@@

Gel-based active electrodes (g.LADYbird, g.Tec, Austria) are used on the same locations Fz, Cz, Pz, P3,P4, Oz, PO7 and PO8.  Reference is set to the right ear lobe and ground is preset as the AFz position.   Sampling frequency is slightly different, and is set to 250 Hz, which is the closest possible to the one used with the other dataset.

@@

The patch size is $S_P = 4.3s \times 4.3s$ pixels, where $s$ is the scale of the local patch and it is an input parameter of the algorithm. The P300 event can have a span of $400$ ms and its amplitude can reach $ 10 \mu V $~\citep{Rao2013}.  Hence it is necessary to utilize a size patch $S_P$ that could capture an entire transient event. With this purpose in consideration, the $s$ value election is essential.

@@

\noindent where $ \lambda $ is the length in seconds covered by the patch, $ Fs $ is the sampling frequency of the EEG signal (downsampled to 16 Hz) and  $\Delta  \mu V $ corresponds to the amplitude in microvolts that can be covered by the height of the patch. The geometric structure of the patch forces a squared configuration, then we discerned that by using $ s =s_x =s_y = 3 $ and $ \gamma = 4 $  the local patch and the descriptor can identify events of 9 $ \mu V $ of amplitude, with a span of $ \lambda = 0.56$ seconds.  This also provides that $ 1 $ pixel represents $ \frac{1}{\gamma}= \frac{1}{4} \mu V $ on the vertical direction and $\frac{1}{F_s.\gamma}=\frac{1}{64}$ seconds on the horizontal direction. Descriptors  $\mathbf{kp}$  are located at $ (x_{kp}, y_{kp} )= ( 0.55 Fs.\gamma, z(c) )= (35,  z(c)) $ for the corresponding channel $c$ (see Eq.~\ref{eq:zerolevel}).   In this way the whole transient event is captured. 
Figure~\ref{fig:sampledescriptor2} shows a patch of a signal plot covering the complete amplitude (vertical direction) and the complete span of the signal event (horizontal direction).

@@

>>>>

% Keywords
\keyword{EEG; BCI; SIFT; P300; ALS; NBNN; HGOSP}

@@

\begin{equation}
z(c) = \left \lfloor{ \frac{\max_{1 \leq n\leq n_{max}} \tilde{x}(n,c)  - \min_{1 \leq n\leq n_{max}} \tilde{x}(n,c) }{2} }\right \rfloor -   \left \lfloor{ \frac{\max_{1 \leq n\leq n_{max}} \tilde{x}(n,c)  + \min_{1 \leq n\leq n_{max}} \tilde{x}(n,c)}{ 2} }\right \rfloor
\label{eq:zerolevel}
\end{equation}

@@

On the generated image $I$, a keypoint $\mathbf{kp}$ is placed on a pixel $(x_{kp}, y_{kp})$ over the image plot and a window around the keypoint is considered. A local image patch of size $S_p \times S_p$ pixels is constructed by dividing the window in $16$ blocks of size $3s$ each one,  where $s$ is the scale of the local patch and it is an input parameter of the algorithm. It is arranged in a $4 \times 4$ grid and the pixel $ \mathbf{kp}$ is the patch center, thus $S_p = 4.3.s $ pixels.

@@

\noindent  where $\mathbf{p}$ is a pixel from  the $i,j$-block $B_{i,j}$, $ i,j \in \{1,2,3,4\} $,  $\theta$ is the angle bin with $ \theta \in \{0, 45, 90, 135, 180, 225, 270, 315\} $,  $ |J(\mathbf{p})| $ is the norm of the gradient vector in the pixel $\mathbf{p}$ and it is computed using finite differences, $\angle J(\mathbf{p}) $ is the angle of the gradient vector, and $ w_\mathrm{ang}(\cdot) $  and $ w_{ij}(\cdot) $ are linear interpolation functions used by Lowe and Vedaldi et al. in~\citep{Lowe2004,Vedaldi2010}.  Lastly, the fixed value of $ 3 $ is a magnification factor which corresponds to the number of pixels per each block when $s = 1$.  As the patch has  $16$ blocks and  $8$ bin angles are considered, a descriptor of $128$ dimension is obtained. It can be observed that the histogram is computed by multiplying by $ |J(\mathbf{p})| $, so the method considers both, the magnitude and the orientation of the gradient vector.

%\begin{figure}[htp]
%\centering
%\includegraphics[width=8cm]{histogramchart.eps}
%\includegraphics[width=6cm]{samplegradients.png}
%\caption{Single trial segments $S_i$ are averaged for the 6 rows and 6 columns. From the averaged signal, the image of the signal plot is generated and each descriptor is computed.  By comparing each descriptor against the set of templates, the P300 ERP can be detected, and finally the desired letter from the matrix can be inferred.}
%\label{fig:fdsfdsfsd}
%\end{figure}
%
%\begin{figure}[htb]
%\centering
%\subfigure[Plot of the signal, a keypoint and the corresponding patch.]{\includegraphics[width=.49\linewidth]{histogramchart.eps}\label{histogram}}
%\subfigure[Orientations on two blocks of the patch.]{\includegraphics[width=.49\linewidth]{samplegradients.png}\label{samplegradients}}
%\caption{ Example of a patch and a scheme of the orientation's histogram computation.}
%\label{fig:histogramchart}
%\end{figure}

@@

\begin{enumerate}
\item \label{paso1}Highlight randomly the rows and columns from the matrix.  There is one row and one column that should match the letter selected by the subject.
\item  \label{paso2} Repeat step~\ref{paso1} $k_a$ times, obtaining the single trial segments $S_1(n,c),\dots,S_{k_a}(n,c)$, of the EEG signal where the variables $n \in \{1, \dots, n_{max}\}$ and $c \in \{1,2,\dots,Ch\}$ correspond to sample points and channel, respectively. The parameter $k_a$ is the number of repetitions and it is an input parameter of the algorithm.
\item \label{paso3} Compute the Ensemble Average by
\begin{equation}
x(n,c)= \frac{1}{k_a}\sum_{i=1}^{k_a}S_i(n,c),n \in \{1, \dots, n_{max}\}, c \in \{1,\dots,Ch\}
\label{averaging}
\end{equation}  
for each row and column.

\item \label{paso4}Plot the signals $x(n,c)$,  $n \in \{1, \dots, n_{max}\}$, $c \in \{1,\dots,Ch\}$,  according Section~\ref{Plot}.

@@

where $NN_T(d^l_u), \;l\in\{row,{col}\}$ is the set of the $k$ nearest neighbors to $d^l_u$ and $q$ is a template descriptor that belongs to $NN_T(d^l_u)$.  This set is obtained according to the k-NBNN  algorithm~\citep{Boiman2008}.

@@

Gel-based active electrodes (g.LADYbird, g.Tec, Austria) are used on the same locations Fz, Cz, Pz, Oz, P3,P4, PO7 and PO8.  Reference is set to the right ear lobe and ground is preset as the AFz position.   Sampling frequency is slightly different, and is set to 250 Hz, which is the closest possible to the one used with the other dataset.

@@

The patch size is $S_P = 4.3.s \times 4.3.s$ pixels, where $s$ is the scale of the local patch and it is an input parameter of the algorithm. The P300 event can have a span of $400$ ms and its amplitude can reach $ 10 \mu V $~\citep{Rao2013}.  Hence it is necessary to utilize a size patch $S_P$ that could capture an entire transient event. With this purpose in consideration, the $s$ value election is essential.

@@

\noindent where $ \lambda $ is the length in seconds covered by the patch, $ Fs $ is the sampling frequency of the EEG signal (downsampled to 16 Hz) and  $\Delta  \mu V $ corresponds to the amplitude in microvolts that can be covered by the height of the patch. The geometric structure of the patch forces a squared configuration, then we discerned that by using $ s =s_x =s_y = 3 $ and $ \gamma = 4 $,  the local patch and the descriptor can identify events of 9 $ \mu V $ of amplitude, with a span of $ \lambda = 0.56$ seconds.  This also provides that $ 1 $ pixel represents $ \frac{1}{\gamma}= \frac{1}{4} \mu V $ on the vertical direction and $\frac{1}{F_s.\gamma}=\frac{1}{64}$ seconds on the horizontal direction. Descriptors  $\mathbf{kp}$  are located at $ (x_{kp}, y_{kp} )= ( 0.55 Fs.\gamma, z(c) )= (35,  z(c)) $ for the corresponding channel $c$ (see Eq.~\ref{eq:zerolevel}).   In this way the whole transient event is captured. 
Figure~\ref{fig:sampledescriptor2} shows a patch of a signal plot covering the complete amplitude (vertical direction) and the complete span of the signal event (horizontal direction).

@@

El uso de descriptores basados en histogramas, presentado en este trabajo, también puede ser usado para derivar una métrica acerca de la forma de la señal de P300.
